---
title: "extendedFamily"
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
<!-- badges: start -->
[![R-CMD-check](https://github.com/gmcmacran/extendedFamily/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gmcmacran/extendedFamily/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/gmcmacran/extendedFamily/branch/main/graph/badge.svg)](https://app.codecov.io/gh/gmcmacran/extendedFamily?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/extendedFamily)](https://CRAN.R-project.org/package=extendedFamily)
<!-- badges: end -->

extendedFamily adds new links to R's generalized linear models. These families are drop in additions to existing families.

Links:

* loglog
* logc
* identity
* odds-power

## Logit and Loglog: Mathematical Comparison

For the binomial family, the link is usually the logit but there are other options. The loglog model assigns a lower probability for X ranging from -5 to 2. For X over 2, the models are essentially indistinguishable. This can lead to improved performance when the response rate is much lower than 50%.

```{r graphExample, message=FALSE, echo = FALSE}
library(stats)
library(dplyr)
library(extendedFamily)
library(ggplot2)
library(yardstick)


logitLink <- binomial(link = "logit")
loglogLink <- binomialEF(link = "loglog")

temp <- bind_rows(
  tibble(X = seq(-5, 5, .01)) %>%
    mutate(
      Probability = logitLink$linkinv(X),
      Link = "logit"
    ),
  tibble(X = seq(-5, 5, .01)) %>%
    mutate(
      Probability = loglogLink$linkinv(X),
      Link = "loglog"
    )
)

ggplot(temp, aes(x = X, y = Probability, colour = Link)) +
  geom_line() +
  scale_x_continuous(breaks = seq(-5, 5, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .1))
```


## Logit and Loglog: Model Performance on Real World Data

The heart data contains info on 4,483 heart attack victims. The goal is to predict if a patient died in the next 48 hours following a myocardial infarction. The low death rate makes this dataset a good candidate for the loglog link.

```{r dataExample, message=FALSE}
data(heart)

heart %>%
  summarise(deathRate = mean(death))
```

Only the family object needs to change to use the loglog link. 

```{r trainModels}
glmLogit <- glm(
  formula = death ~ anterior + hcabg + kk2 + kk3 + kk4 + age2 + age3 + age4,
  data = heart, family = binomial(link = "logit")
)
glmLoglog <- glm(
  formula = death ~ anterior + hcabg + kk2 + kk3 + kk4 + age2 + age3 + age4,
  data = heart, family = binomialEF(link = "loglog")
)
```

AUC improved by changing the link.

```{r calcAUC}
predictions <- heart %>%
  select(death) %>%
  mutate(
    death = factor(death, levels = c("0", "1")),
    logitProb = predict(object = glmLogit, newdata = heart, type = "response"),
    loglogProb = predict(object = glmLoglog, newdata = heart, type = "response")
  )

roc_auc(data = predictions, truth = death, event_level = "second", logitProb)

roc_auc(data = predictions, truth = death, event_level = "second", loglogProb)
```

## Tidymodels

The family objects integrate with Tidymodels.

```{r tidy, message=FALSE}
library(tidymodels)

heart <- heart %>%
  mutate(death = factor(death, levels = c("0", "1")))

parsnip_fit <-
  logistic_reg() %>%
  set_engine("glm", family = binomialEF("loglog")) %>%
  fit(death ~ anterior + hcabg + kk2 + kk3 + kk4 + age2 + age3 + age4, data = heart)

testPredictions <- parsnip_fit %>%
  predict(new_data = heart, type = "prob")
testPredictions <- heart %>%
  select(death) %>%
  bind_cols(testPredictions)

testPredictions %>%
  roc_auc(truth = death, event_level = "second", .pred_1)
```
